Leadership across cultures and contexts: an exploratory analysis

### 2020-05-02 leadership_across_cultures_contexts.Rmd
Heatmaps. costs and benefits.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=F)
library(readr)
library(knitr)
library(pander)
library(patchwork)
library(gt)
loadd(all_study_vars)
loadd(df_qual)
loadd(df_func)
loadd(plot_func_qual)
loadd(qualities_support_txt)
loadd(qualities_support_cult)
loadd(functions_support_txt)
loadd(functions_support_cult)
loadd(highstatus_plot)
loadd(plot_costs_benefits)
loadd(plot_pubdate)
loadd(m_lpca_qual)
loadd(m_lpca_func)
loadd(multi_allvars)
loadd(p_heatmap_subsis)
loadd(p_heatmap_region)
loadd(p_heatmap_sex)
loadd(p_heatmap_groups)
source("R/cheapobjects.R")
```

@@

We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level. Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record. We computed the proportion of text records supporting each variable, and their 95% confidence intervals, using logistic generalized linear mixed effects models with random intercepts for document authors nested within cultures (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or theoretical framework), controlling for one potential source of bias as a fixed effect: the date of publication. For each variable, we then computed the proportion of cultures with at least one text record with evidence for that variable, with confidence intervals computed using a cluster bootstrap.

@@

@@

```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of group type, subsistence strategy, continental region, and leader sex, with random effects for ethnographer nested within culture. Values are on the response scale (probability)."}
((plot_emm_status_group2 / plot_emm_status_sex2) | (plot_emm_status_subsistence2 / plot_emm_status_region2)) + plot_annotation(title = "Predictors of high status", tag_levels = 'A')
```

@@

We then explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, and levels of group context. In general, leaders at the highest supracommunity political level had the most evidence for receiving benefits and the least evidence for suffering costs and leaders of military groups has the highest total costs. See Figure \@ref(fig:leadertotalbenefitscosts).

```{r leadertotalbenefitscosts, fig.cap="Estimated marginal means of A. leader total benefits, and B. total costs, as a function of group structure, based on a logistic mixed effects regression model with random intercepts for ethnographer nested within culture. Values are on the response scale."}
p1 <- 
  hagenutils::ggemmeans(emmeans(m_ldrtotben, type = 'response', specs = 'group.structure2')) +
  scale_x_continuous(limits = c(0, 0.1)) +
  labs(title = "Total leader benefits", x = "", y = "")

p2 <- hagenutils::ggemmeans(emmeans(m_ldrtotcost, type = 'response', specs = 'group.structure2')) +
  scale_x_continuous(limits = c(0, 0.1)) +
  labs(title = "Total leader costs", x = "\nProbability", y = "")

p1 + p2 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')

@@

## Effects of subsistence strategy and region on function and quality variables

@@

plot_morality_region / plot_prosocial_region
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=F)
library(readr)
library(knitr)
library(pander)
library(patchwork)
library(gt)
loadd(all_study_vars)
loadd(df_qual)
loadd(df_func)
loadd(plot_func_qual)
loadd(qualities_support_txt)
loadd(qualities_support_cult)
loadd(functions_support_txt)
loadd(functions_support_cult)
loadd(highstatus_plot)
loadd(plot_costs_benefits)
loadd(plot_pubdate)
loadd(m_lpca_qual)
loadd(m_lpca_func)
loadd(multi_allvars)
loadd(multi_aic)
loadd(p_heatmap_subsis)
loadd(p_heatmap_region)
loadd(p_heatmap_sex)
loadd(p_heatmap_groups)
source("R/cheapobjects.R")
```

@@

We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level. Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record. We computed the proportion of text records supporting each variable, and their 95% confidence intervals, using logistic generalized linear mixed effects models with random intercepts for document authors nested within cultures (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or theoretical framework). For each variable, we then computed the proportion of cultures with at least one text record with evidence for that variable, with confidence intervals computed using a cluster bootstrap.

@@

## Variation in study variables by subsistence strategy, continental region, leader sex, and group structure

To investigate variation in functions, qualities, costs, and benefits, we fit a multiple logistic mixed effects regression model of each variable as a function of subsistence strategy, continental region, leader sex, and group structure, with random intercepts for author nested within culture. We then compared the AIC value of each model to a model with only the fixed and random intercepts, retaining `r nrow(multi_aic)` models that showed at least a two point reduction in AIC value. We then created heatmaps that displayed the relative estimated marginal means for each level of these categorical explanatory variables for each outcome variable. See Figures \@ref(fig:heatmapsubsis), \@ref(fig:heatmapregion), \@ref(fig:heatmapsex), and \@ref(fig:heatmapgroups), and Tables \@ref(tab:variablevars) and \@ref(tab:universalvars).

For subsistence strategy, hunter-gatherers had relatively more evidence for polygyny, mating benefits, no coercive authority, political appointments and military command than other subsistence strategies. Pastoralists had more evidence for ritual functions, supernatural, aggressiveness and high quality spouse. Hunter-gatherers and pastoralists both had more evidence for high status, providing counsel/direction, and providing subsistence. The remaining subsistence strategies generally had relatively less evidence for all variables, except that the mixed strategy had relatively more evidence for organizing cooperation, distributing resources, and prosocial investment. See Figure \@ref(fig:heatmapsubsis).

For continental region, there was little strong patterning, except that Africa generally had relatively less evidence for any retained variable except providing counsel/direction, and South America had relatively more evidence for several variables, including age, social services, and high status. See Figure \@ref(fig:heatmapregion).

There was strong patterning of relative evidence for almost all variables by leader sex. Female leaders had relatively more evidence for high quality spouse, mating benefit, polygyny, high status, and age. Male leaders had relatively more evidence for social services, political appointments, prosocial investment, and loss of social services. Records pertaining to both sexes had relatively more evidence for organizing cooperation, providing subsistence, ritual functions, bravery, experience/accompished, and few that overlapped with males, such as military command and distributing resources. See Figure \@ref(fig:heatmapsex).

There was also strong patterning of relatively evidence for almost all variables by group structure. There was relatively more evidence for military traits in military groups, relatively more evidence for ritual and supernatural traits in relgious groups, relatively more evidenc for organizing cooperation and providing subsistence in economic groups, and so forth. Kin groups had relatively more evidence for providing counsel/direction, prosocial investment, age, and high status. Community and residential groups had relatively low evidence for most variables, except that there was relatively more evidence for high quality spouses among residential group leaders. See Figure \@ref(fig:heatmapgroups).

## Total leader benefits and costs

We then explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, and levels of group context. In general, leaders at the highest supracommunity political level had the most evidence for receiving benefits and the least evidence for suffering costs and leaders of military groups has the highest total costs. See Figure \@ref(fig:leadertotalbenefits).

```{r leadertotalbenefits, fig.cap="Estimated marginal means of A. leader total benefits, and B. total costs, as a function of group structure, based on a logistic mixed effects regression model with random intercepts for ethnographer nested within culture. Values are on the response scale."}
# p1 <- 
#   hagenutils::ggemmeans(emmeans(m_ldrtotben, type = 'response', specs = 'group.structure2')) +
#   scale_x_continuous(limits = c(0, 0.1)) +
#   labs(title = "Total leader benefits", x = "", y = "")
# 
# p2 <- hagenutils::ggemmeans(emmeans(m_ldrtotcost, type = 'response', specs = 'group.structure2')) +
#   scale_x_continuous(limits = c(0, 0.1)) +
#   labs(title = "Total leader costs", x = "\nProbability", y = "")
# 
# p1 + p2 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')

theplots <- all_emms(m_ldrtotben, c('subsistence', 'group.structure2', 'region', 'demo_sex'), 0.1, 'Predictors of leader benefits')

((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = 'Predictors of leader benefits', tag_levels = 'A')

```{r leadertotalcosts}
theplots <- all_emms(m_ldrtotcost, c('subsistence', 'group.structure2'), 0.06, 'Predictors of leader costs')

theplots[[1]] / theplots[[2]] + plot_annotation(title = 'Predictors of leader costs', tag_levels = 'A')

@@

```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of group type, subsistence strategy, continental region, and leader sex, with random effects for ethnographer nested within culture. Values are on the response scale (probability)."}
# ((plot_emm_status_group2 / plot_emm_status_sex2) | (plot_emm_status_subsistence2 / plot_emm_status_region2)) + plot_annotation(title = "Predictors of high status", tag_levels = 'A')

theplot <- all_emms(m_status_group2, c('subsistence', 'group.structure2', 'region', 'demo_sex'), 0.8, 'Predictors of leader status')

((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = title, tag_levels = 'A')
```

@@

@@

```{r variablevars}

multi_aic %>% 
  select(
    var,
    Variable, 
    Sum, 
    uniAIC, 
    multiAIC, 
    AIC_diff,
    pvalue_subsis,
    pvalue_region,
    pvalue_female,
    pvalue_groups
    ) %>% 
  mutate(
    var = map_chr(var, ~ str_split(., '_')[[1]][[1]])
  ) %>% 
  arrange(Variable) %>% 
  kable(
  caption = 'Retained logistic mixed effects models of each variable as functions of subsistence strategy, continental region, leader sex, and group stuctures. Models were retained due to improvement in AIC values relative to an intercept-only model. Sum: the number of rows with evidence for the variable. uniAIC: AIC of intercept-only model. multiAIC: AIC of multivariate model. AIC_diff: difference in AIC values. P-values are the unadjusted p-values of each explanatory variable in the model.',
  digits = 2
  )
```

```{r universalvars}
multi_allvars %>%
  dplyr::filter(AIC_diff >= -2) %>% 
  select(
    var,
    Variable, 
    Sum, 
    uniAIC, 
    multiAIC, 
    AIC_diff,
    pvalue_subsis,
    pvalue_region,
    pvalue_female,
    pvalue_groups
    ) %>% 
  mutate(
    var = map_chr(var, ~ str_split(., '_')[[1]][[1]])
  ) %>% 
  arrange(Variable) %>% 
  kable(
  caption = 'Logistic mixed effects models that did not show improvement in AIC values relative to an intercept-only model. These study variables either did not vary much by subsistence strategy, continental region, leader sex, or group stuctures, or had little evidence. Sum: the number of rows with evidence for the variable. uniAIC: AIC of intercept-only model. multiAIC: AIC of multivariate model. AIC_diff: difference in AIC values. P-values are the unadjusted p-values of each explanatory variable in the model.',
  digits = 2
  )

<!-- ## Effects of subsistence strategy and region on function and quality variables -->

@@

# plot_morality_region / plot_prosocial_region
```
